Crossing Origins by Crossing Formats

### 2012-11-30 main.tex
intro

@@

%3 pages
\section{Crossing origins by crossing formats}
\label{sec:cross}
% Describe how the attacks lead to Same-Origin Policy bypass.

This section describes how formats can be crossed and how that can be abused 
to cross origins by circumventing the Same-Origin Policy. We begin by 
generalizing the problem of crossing formats to \emph{polyglots} and then describe the  
characteristics of a malicious polyglot. From that we derive two attack vectors and 
show how previous work on the subject relate to these vectors.

@@

By crossing origins we mean being able to request and access content
which is normally restricted under the Same-Origin Policy. 
Recall that SOP dictates that a two document can only access each others 
content and resources if they share the same origin. Similarly this applies to communication as 
well; a document can only directly communicate with the server from which it originated. 
This is not to be confused with Cross-Origin Resource Sharing (CORS)~\cite{w3-cors}, which is an 
intentional relaxing of the SOP.

While there are exceptions to this policy, e.g. images, scripts, and style sheets 
are allowed to be included as resources across origins, access to these resources is 
carefully restricted to prevent information leaks. As an example, the including 
document is prevented from accessing the image data of images loaded across 
origins.

Not all elements are as carefully restricted from information leaks as images. 
Scripts loaded across origins become part of the document and inherit the 
origin of the including document, which allows the script to communicate with the 
server from which the \emph{document} originated. Such scripts are also able to interact 
with the document, e.g. by adding nodes, which in turn require new content to be 
requested. Since these requests are not restricted by the SOP, this creates a side 
channel that permits cross-origin information leakage. At the same time, the including document
is prevented from inspecting the source of the script and can only observe the public 
side-effects that the script produce as it is executed.

Other examples of problematic elements include the object- and embed-tags. These 
tags allow inclusion of resources that might require a 
plug-in to run. As with images, the content handled by the plug-in retains its origin, 
but the browsers are forced to rely on the plug-ins to employ correct security 
measures to prevent information leaks. The plug-in is selected based on the MIME-type of the content, but
because the server delivering the content might not be able to determine its
format, the object tag allows a developer to set the type-attribute to guide the
browser in which plug-in to run. When the type attribute is used, the
corresponding plug-in is run regardless of which MIME-type the server returns
and it is up to the plug-in to respond to and handle the content it is served. While the containing page is restricted 
from directly accessing the content handled by the plug-in, a number of plugins provide 
an API for interacting with the plug-in and the data. As we will show, this can lead to 
to cross-origin information leaks.

@@

A polyglot is defined as ``A person who knows several languages.'', and the term
is widely used in various fields of science. In computer science, a \emph{polyglot} is
a program that is valid in multiple programming languages. In this paper we use
a broader definition of a programming language, not limited to code meant for
compilation to machine code or scripting languages, but extended to any format
that requires interpretation before rendering.

@@

Certain languages are particularly suitable for creating polyglots. These
languages either have a lot of constructs in common with other languages, such
as \emph{C}, or are defined such that the parser ignores that which it cannot
interpret, such as \emph{HTML}. The latter allows for ample opportunity to hide
any code specific to another language, as long as there is no overlap with the
syntax of the first language.

@@

In a web context, coercing content to be interpreted as a different type be 
done easily. If the content is loaded using an image-tag it will be interpreted 
as an image, if a script-tag is used it will be interpreted as a script. Some 
tags even let the developer decide which type to interpret the content as. 
To prevent abuse, the browsers employ content-type 
sniffing for certain content-types, a practice which has historically led to 
security issues.
\todo{explain the issues and cite Barthe et al.}

@@

We have identified two vectors in which malicious polyglots can be used for 
exploitation; syntax injection, and content repurposing.

@@

In a cross-site scripting attack, HTML syntax is injected into a preexisting
HTML document via vulnerable user supplied parameters. Similarly, a syntax
injection attack consists of injecting foreign syntax into preexisting content
of a different format, thereby creating a polyglot. The polyglot content can
then be interpreted as either format; the benign original format, or the
malicious user-supplied format.

@@

\todo{diagram}
The target in this scenario is a user of the vulnerable website. The polyglot is
served in the benign format on the vulnerable server. While the user is logged in to
the service that the vulnerable website provides, the user unsuspectingly visits
the attackers malicious website. The page embeds the polyglot content, but
coerces the browser to interpret the content as the malicious format. 
The capabilities of the malicious format determine the severity of the consequences.
The attacker can
exploit the vulnerability, either by making requests to the vulnerable website
in the context of the user; or by extracting information about the user that is
stored on the vulnerable website.

@@

%The same-origin policy is intended to prevent a document from one origin from
%accessing the content of a document of another origin. 
As an example the
same-origin policy allows a script-tag in one document to include JavaScript of
another origin and execute it in the context of that document. If an attacker is
able to inject an HTML document with JavaScript syntax in such a way that it
became valid JavaScript, the attacker can include the document across origins
via a script-tag and extract its content, thereby circumventing the same-origin
policy.

@@

\todo{Mention and explain known attack instances}

\subsection{Content repurposing}
\todo{diagram}
% Description of attack technique 
A content repurposing attack targets the users of a website
that allows and serves user supplied files. Examples are a file sharing
service, an image database, or a job search service. Such a service 
accepts a limited set of file-types, and the files are verified to
be of one of these types, e.g. an image, before being served to the end
user. By submitting a polyglot to such a service, an attacker can trick the
service into serving malicious content under the origin of the service. The
service will verify the file, assuming that it is of the benign format, but once
the content is served an attacker can coerce the content to be interpreted as
the malicious format.

@@

% Scenarios
\todo{Mention and explain known attack instances}

@@

In this section we give concrete examples of the attacks described in 
Section~\ref{sec:cross}, using the PDF format as the running example. We begin
by detailing how the design decision made in the PDF-standard make it ideal 
for creating malicious polyglots.

@@

When embedded in a web page, many of the PDF communication methods utilize
the underlying browser to make requests by navigating the browser to the target 
URL. This will unload the current page in which the document is embedded. Page 
navigation is allowed wtih respect to the SOP, since the response is not visible 
to the document. Such methods can be used in CSRF attacks, but would 
defeat the purpose of requesting information as the document cannot process the 
response.

@@

\paragraph{Embedded Flash}
Cross-origin communication in Flash is restricted by the Same-Origin Policy,
but also support cross-origin communication via Cross-Origin Resource Sharing 
(CORS). Using CORS, a web server can relax the SOP to allow access to specified content.
As mentioned, the embedded Flash file inherits the origin of the document.
However, this is not a restriction, since an attacker can set a cross-origin policy 
that allow the request. This allows for a two way communication that is not 
visible to the victim.

@@

In the case of PDF-based polyglots, the browser can intervene when there is a 
mismatch between the content-type provided by the server and the type attribute 
of the object tag. By acknowledging that there can be a potential security issue in 
interpreting the supplied content in the requested format and alerting the user to this
threat, the vulnerability can be mitigated.

@@

>>>>

\todo{Say why sanitization and CSP are of no use}

\todo{Mention GIFAR and perhaps other profile examples and say we're
  the first to perform a systematic study}

@@

%3 pages
\section{Crossing origins by crossing formats}
\label{sec:cross}
% Describe how the attacks lead to same-origin policy bypass.

This section describes how formats can be crossed and how that can be abused 
to cross origins by circumventing the same-origin policy. We begin by 
generalizing the problem of crossing formats to \emph{polyglots} and then describe the  
characteristics of a malicious polyglot. From that we derive two attack vectors and 
show how previous work on the subject relate to these vectors.

@@

By crossing origins we mean being able to request and access content
which is normally restricted under the same-origin policy. 
Recall that SOP only allows two documents accessing each others'
content and resources if they share the origin. Similarly,
a document can only directly communicate with the server from which it originated. 
This is not to be confused with \emph{cross-origin resource sharing} (CORS)~\cite{w3-cors}, which is an 
intentional relaxation of SOP.

While there are exceptions to this policy, e.g. images, scripts, and style sheets 
are allowed to be included as resources across origins, access to these resources is 
restricted to prevent information leaks. As an example, the including 
document is prevented from accessing the image data of images loaded across 
origins.

Not all elements are as carefully restricted from information leaks as images. 
Scripts loaded across origins become part of the document and inherit the 
origin of the including document, which allows the script to communicate with the 
server from which the \emph{document} originated. Such scripts are also able to interact 
with the document, e.g. by adding nodes, which in turn require new content to be 
requested. Since these requests are not restricted by SOP, this creates a side 
channel that permits cross-origin information leakage. At the same time, the including document
is prevented from inspecting the source of the script and can only observe the public 
side-effects that the script produce as it is executed.

Other examples of problematic elements include the \emph{object} and
\emph{embed} tags. These 
tags allow inclusion of resources that might require a 
plug-in to run. As with images, the content handled by the plug-in retains its origin, 
but the browsers are forced to rely on the plug-ins to employ correct security 
measures to prevent information leaks. The plug-in is selected based on the MIME-type of the content, but
because the server delivering the content might not be able to determine its
format, the \emph{object} tag allows a developer to set the type-attribute to guide the
browser in which plug-in to run. When the type attribute is used, the
corresponding plug-in is run regardless of which MIME-type the server returns
and it is up to the plug-in to respond to and handle the content it is served. While the containing page is restricted 
from directly accessing the content handled by the plug-in, a number of plugins provide 
an API for interacting with the plug-in and the data. As we will show, this can lead to 
to cross-origin information leaks.

@@

A polyglot is perhaps most commonly known as a person who speaks several
languages. However, the term
is widely used in several scientific fields. In computer science, a \emph{polyglot} is
a program that is valid in multiple programming languages. In this paper we use
a broader definition of a programming language, not limited to code meant for
compilation to machine code or scripting languages, but extended to any format
that requires interpretation before rendering.

@@

Certain languages are particularly suitable for creating polyglots. These
languages either have a lot of constructs in common with other languages, such
as the C language, or are defined such that the parser ignores that which it cannot
interpret, such as HTML. The latter allows for ample opportunity to hide
any code specific to another language, as long as there is no overlap with the
syntax of the first language.

@@

Coercing content to be interpreted
as a different type can be accommodated in the context of the web. 
If the content is loaded using an \emph{image} tag it will be interpreted 
as an image, if a \emph{script} tag is used it will be interpreted as a script. Some 
tags even let the developer decide which type to interpret the content as. 
To prevent abuse, the browsers employ content-type 
sniffing for certain content-types, a practice which has historically led to 
security issues.
\todo{explain the issues and cite Barthe et al.}

@@

We have identified two vectors in which malicious polyglots can be used for 
exploitation: syntax injection and content repurposing.

@@

In a cross-site scripting attack, HTML syntax is injected into a preexisting
HTML document via vulnerable user supplied parameters. Similarly, a syntax
injection attack consists of injecting foreign syntax into preexisting content
of a different format, thereby creating a polyglot. The polyglot content can
then be interpreted as either the benign original format or the
malicious user-supplied format.

@@

\todo{diagram with attack scenario. for example: with example domains attacker.com and
  vulnerable.com leaking secret file secret.txt}
The target in this scenario is a user of the vulnerable website. The polyglot is
served in the benign format on the vulnerable server. While the user is logged in to
the service that the vulnerable website provides, the user unsuspectingly visits
the attackers malicious website. The page embeds the polyglot content, but
coerces the browser to interpret the content as the malicious format. 
The capabilities of the malicious format determine the severity of the consequences.
The attacker can
exploit the vulnerability, either by making requests to the vulnerable website
in the context of the user; or by extracting information about the user that is
stored on the vulnerable website.

@@

%The same-origin policy is intended to prevent a document from one origin from
%accessing the content of a document of another origin. 
As an example the
same-origin policy allows a \emph{script} tag in one document to include JavaScript of
another origin and execute it in the context of that document. If an attacker is
able to inject an HTML document with JavaScript syntax in such a way that it
became valid JavaScript, the attacker can include the document across origins
via a \emph{script} tag and extract its content, thereby circumventing the same-origin
policy.

@@

\todo{
Mention and explain known attack instances.
To what extent known attacks such as GIFAR follow this. if not
  we should still mention them}

\subsection{Content repurposing}
\todo{diagram with attack scenario. for example: with example domains attacker.com and
  vulnerable.com leaking secret file secret.txt}
% Description of attack technique 
A content repurposing attack targets the users of a website
that allows and serves user supplied files. Examples are a file sharing
service, an image database, or a job search service. Such a service 
accepts a limited set of file-types, and the files are verified to
be of one of these types, e.g. an image, before being served to the end
user. By submitting a polyglot to such a service, an attacker can trick the
service into serving malicious content under the origin of the service. The
service will verify the file, assuming that it is of the benign format, but once
the content is served an attacker can coerce the content to be interpreted as
the malicious format.

@@

% Scenarios
\todo{
Mention and explain known attack instances.
To what extent known attacks such as GIFAR follow this. if not
  we should still mention them}

@@

In this section we give concrete examples of the attacks described in 
Section~\ref{sec:cross}, using the PDF format as the running example. We begin
by detailing how the design decision made in the PDF-standard make it 
highly suitable
for creating malicious polyglots.

@@

When embedded in a web page, many of the PDF communication methods utilize
the underlying browser to make requests by navigating the browser to the target 
URL. This will unload the current page in which the document is embedded. Page 
navigation is allowed wtih respect to SOP, since the response is not visible 
to the document. Such methods can be used in CSRF attacks, but would 
defeat the purpose of requesting information as the document cannot process the 
response.

@@

\paragraph{Embedded Flash}
Cross-origin communication in Flash is restricted by the same-origin policy,
but also support cross-origin communication via Cross-Origin Resource Sharing 
(CORS). Using CORS, a web server can relax SOP to allow access to specified content.
As mentioned, the embedded Flash file inherits the origin of the document.
However, this is not a restriction, since an attacker can set a cross-origin policy 
that allow the request. This allows for a two way communication that is not 
visible to the victim.

@@

In the case of PDF-based polyglots, the browser can intervene when there is a 
mismatch between the content-type provided by the server and the type attribute 
of the \emph{object} tag. By acknowledging that there can be a potential security issue in 
interpreting the supplied content in the requested format and alerting the user to this
threat, the vulnerability can be mitigated.

@@

Billy's and Jonas' anonimized demos?
